Systemix MAG-SMAG Profiling

### 2024-07-26 R/paper.Rmd
revised manuscript & figures

# Inspect completeness and contamination per sample, then select one
filter(mags_qc, grepl("\\-S15\\.", bin)) %>% dplyr::select(2:3) %>% summary()
mag_samples <- filter(gtdb_taxonomy, grepl("\\-S15\\.", genome))

@@

# Merge metadata and save to file
mags_metadata <- merge(mag_samples, smag_data, all = TRUE)
if (file.exists("4-mags-smag/mags_metadata.tsv") == FALSE) {
  write.table(mags_metadata, "4-mags-smag/mags_metadata.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
}

# Save parsed SMAG metadata
if (file.exists("4-mags-smag/smag_data.tsv") == FALSE) {
  write.table(smag_data, "4-mags-smag/smag_data.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
}

# Save biomining metadata (S15)
if (file.exists("4-mags-smag/mags_biomining.tsv") == FALSE) {
  write.table(mag_samples, "4-mags-smag/mags_biomining.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
}

@@

# Select top phyla and plot a histogram
mags_md_top_phy <- table(mags_metadata_bac$phylum) %>%
  sort(TRUE) %>%
  head(n = 40) %>%
  as.data.frame() %>%
  dplyr::rename(phy_top = Var1, freq = Freq)

plot_depth_top_phyla <- mags_md_top_phy %>%
  ggplot(aes(x = freq, y = reorder(phy_top, freq))) +
  geom_bar(stat = "identity", fill = "#459682") +
  geom_text(aes(label = freq), vjust = 0.5, hjust = -0.25, size = 3) +
  labs(x = "Number", y = "Phylum") +
  scale_x_continuous(limits = c(0, 10000)) +
  theme_classic() +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16))
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_top_phyla.pdf", plot_depth_top_phyla, height = 7, width = 6)

# Filter metadata..
# Set colors for ecosystem
#mags_colors_eco <- colorRampPalette(brewer.pal(12, "Paired"))(length(unique(mags_metadata_bac$ecosystem)))
#mags_colors_eco <- sample(mags_colors_eco)
mags_colors_eco <- c("#1F78B4", "#FF7F00", "#B2DF8A", "#E31A1C", "#A6CEE3",
                     "#6A3D9A", "#33A02C", "#CAB2D6", "#FDBF6F", "#FB9A99")

# Which bacteria from biomining sites are annotated as known?
mags_depth_spec_biom <- mags_metadata_bac %>%
  dplyr::select(mag_name, family, genus, species, ecosystem) %>%
  dplyr::filter(species != "" & ecosystem == "Biomining")

@@

plot_depth_spec <- ggplot(mags_depth_spec) +
  geom_bar(aes(x = species, fill = phylum)) +
  labs(x = "", y = "Number of kSGBs from biomining") +
  scale_fill_manual(values = "#78acc2") +
  coord_flip() +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = c(0.88, 0.12),
        strip.background = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14),
        strip.text.x = element_text(size = 10)) +
  facet_grid(.~ecosystem)
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_biom_species.pdf", plot_depth_spec, height = 3, width = 9)

@@

#add column
mags_depth_spec_eco <- mags_depth_spec_eco %>%
  dplyr::mutate(mags = replace(mags, ecosystem == "Biomining", 1)) %>%
  dplyr::mutate(ratio = count_eco / mags) #max 6.52

plot_depth_spec_eco <- mags_depth_spec_eco %>%
  ggplot(aes(x = reorder(ecosystem, ratio), y = ratio, fill = ecosystem)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(ratio, 2)), vjust = 0.5, hjust = -0.25, size = 3) +
  labs(x = "Ecosystem", y = "Number of kSGBs/MAGs") +
  scale_y_continuous(limits = c(0, 10)) +
  scale_fill_manual(values = mags_colors_eco) +
  coord_flip() +
  theme_classic() +
  theme(legend.position = "none",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_species_eco.pdf", plot_depth_spec_eco, height = 7, width = 6)

@@

plot_depth_sgbs <- mags_depth_sgbs[1:60, ] %>%
  ggplot(aes(x = reorder(phylum, tot), y = tot, fill = cond)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#dd7e8b", "#459682")) +
  scale_y_continuous(limits = c(0, 8000)) +
  labs(x = "Phylum", y = "Number of species") +
  coord_flip() +
  theme_classic() +
  theme(legend.position = c(0.85, 0.15),
        strip.placement = "outside",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))

@@

plot_depth_sgbs_p <- mags_depth_sgbs_p[1:60, ] %>%
  ggplot(aes(x = reorder(phylum, t_phy), y = tot_p, fill = cond_p)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#dd7e8b", "#459682")) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "", y = "% of species") +
  coord_flip() +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))

#merge
plot_depth_sgbs_both <- wrap_plots(plot_depth_sgbs, plot_depth_sgbs_p,
                                   nrow = 1, ncol = 2, widths = c(0.55, 0.5))
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_SGBs_phylum.pdf", plot_depth_sgbs_both, height = 7, width = 7)

# And unknown/known SGBs for biomining?
mags_depth_sgbs_biom <- mags_metadata_bac %>%
  dplyr::filter(ecosystem == "Biomining") %>%
  dplyr::group_by(phylum, species) %>%
  dplyr::summarise(n_spe = n()) %>%
  ungroup() %>%
  dplyr::group_by(phylum) %>%
  dplyr::mutate(t_phy = sum(n_spe)) %>%
  dplyr::filter(species == "") %>%
  dplyr::mutate(diff_k = t_phy - n_spe) %>%
  dplyr::mutate(perc_u = n_spe * 100 / t_phy) %>%
  dplyr::arrange(desc(t_phy)) %>%
  pivot_longer(cols = c(n_spe, diff_k), names_to = "cond", values_to = "tot")

plot_depth_sgbs_biom <- mags_depth_sgbs_biom %>%
  ggplot(aes(x = reorder(phylum, tot), y = tot, fill = cond)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#dd7e8b", "#459682")) +
  scale_y_continuous(limits = c(0, 25)) +
  labs(x = "Phylum", y = "Biomining species") +
  coord_flip() +
  theme_classic() +
  theme(aspect.ratio = 0.875,
        legend.position = c(0.75, 0.3),
        strip.placement = "outside",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_SGBs_phylum_biom.pdf", plot_depth_sgbs_biom, height = 4, width = 6)

@@

plot_depth_usgbs <- mags_depth_usgbs_eco %>%
  ggplot(aes(x = num, y = reorder(phylum, num), fill = ecosystem)) +
  geom_col() +
  #geom_text(aes(label = num), vjust = 0.5, hjust = -0.5, size = 2.5) +
  labs(x = "Number of uSGBs", y = "Phylum") +
  scale_x_continuous(n.breaks = 3) +
  scale_fill_manual(values = mags_colors_eco) +
  theme_classic() +
  theme(legend.position = "none",
        strip.placement = "outside",
        strip.background = element_blank(),
        panel.border = element_blank(),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        strip.text.x = element_text(size = 10)) +
  facet_grid(.~ecosystem, scales = "free")
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_uSGBs_eco.pdf", plot_depth_usgbs, height = 9, width = 14)

# What about the annotated bins at genus level (kGGBs) from biomining?
mags_depth_kggbs_biom <- mags_metadata_bac %>%
  dplyr::select(mag_name, mag_id, phylum, family, genus, ecosystem) %>%
  dplyr::filter(genus != "" & ecosystem == "Biomining") #nrow=21

@@

plot_depth_kggbs_biom <- mags_depth_kggbs_biom %>%
  ggplot(aes(y = reorder(genus, norm), x = norm, fill = ecosystem)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_x_continuous(limits = c(0, 101), n.breaks = 5) +
  scale_fill_manual(values = mags_colors_eco) +
  labs(x = "% kGGBs", y = "Known genus from biomining") +
  theme_classic() +
  theme(legend.title = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_kGGBs_biom.pdf", plot_depth_kggbs_biom, height = 7, width = 6)

### --EXPLORE OTHER TAXA-- ###
##inspect unannotated family-level bins (uFGBs)? #1166
##what about novel bacteria at genus level?
#mags_depth_novel_gen <- mags_metadata_bac %>%
#  dplyr::filter(genus == "") %>%
#  dplyr::group_by(phylum, genus, ecosystem) %>%
#  dplyr::summarise(gen_unk = n()) %>%
#  #dplyr::arrange(desc(gen)) %>%
#  ungroup() %>%
#  dplyr::group_by(genus) %>%
#  dplyr::mutate(tot_gen = sum(gen_unk)) %>%
#  dplyr::mutate(perc = gen_unk * 100 / tot_gen)
#
##get unique classes of unknown genus by ecosystem
##Leptospirillia, Sulfobacillia
#mags_md_unk_gen2 <- mags_metadata_bac %>%
#  dplyr::filter(genus == "") %>%
#  dplyr::group_by(class, ecosystem) %>%
#  tally() %>%
#  #slice(n()) %>%
#  ungroup()
#unique(mags_md_unk_gen2[mags_md_unk_gen2[, "ecosystem"] == "Biomining", "class"])
#
##unique genus from biomining: Acidithrix, Ferrithrix, SCQM01
#unique(mag_samples[mag_samples[, "family"] != "", "genus"])
#mags_metadata_bac %>% filter(grepl("Acidithrix", genus, ignore.case = TRUE) == TRUE)
#
#plot_md_unk_cla <- ggplot(mags_md_unk_gen2[mags_md_unk_gen2$n > 10, ],
#                          aes(x = reorder(class, n), y = n, fill = ecosystem)) +
#  geom_bar(stat = "identity") +
#  labs(x = "Class", y = "Count") +
#  scale_y_continuous(limits = c(0, 250)) +
#  theme_minimal() +
#  coord_flip() +
#  facet_wrap(~ecosystem)
#ggsave("plot_md_unk_class_eco.pdf", plot_md_unk_cla, height = 7, width = 10)
#
##class of interest for acidic soils:
##Alphaproteobacteria, Gammaproteobacteria
##Acidobacteriae, Thermoleophilia, Sulfobacillia, Acidimicrobiia
#table(mags_metadata_bac[mags_metadata_bac[, "phylum"] == "Proteobacteria", "class"])
#table(mags_metadata_bac[mags_metadata_bac[, "class"] == "Acidimicrobiia", "order"])
#
##family of interest reported for Systemix [ref 2022]:
##Pseudomonadaceae (mayor), Flavobacteriaceae, Erwiniaceae
#mags_md_fams_acid <- mags_metadata_bac %>%
#  dplyr::filter(family == "Pseudomonadaceae") %>%
#  dplyr::group_by(ecosystem) %>%
#  dplyr::summarise(count = n())
```

@@

# Select S15 sample
mags_qc <- mags_qc %>% filter(grepl("Refined-S15", bin))

@@

# Plot
plot_mags_qc <- mags_qcs %>%
  ggplot(aes(x = mags, y = fct_reorder(label, mags), fill = label)) +
  geom_col(position = "dodge2") +
  #geom_text(aes(label = mags), vjust = 0.5, hjust = -0.25, size = 3) +
  scale_fill_manual(values = c("#db7971", "#5ca650", "#859ac7")) +
  labs(x = "Number of MAGs", y = "") +
  theme_classic() +
  theme(aspect.ratio = 0.875,
        axis.text = element_text(size = 12),
        axis.title.x = element_text(size = 14),
        legend.position = "none")
ggsave("article_profiling/2-statistics/plots/plot_biom_MAGs_QC.pdf", plot_mags_qc, height = 4, width = 6)
```

@@

@@

## Figure 1 & supplementary 1

```{r fig1}
# Merge plots using patchwork
fig_1a <- wrap_plots(plot_tree_s15_md, plot_mags_qc, plot_depth_sgbs_biom,
                     design = "\nAAAB\nAAAC", heights = c(1, 0.85))
#ggsave("article_profiling/fig_1a.pdf", fig_1a, height = 8, width = 15)
fig_1b <- wrap_plots(plot_depth_sgbs_both, plot_depth_spec_eco, plot_depth_kggbs_biom,
                     design = "\nAAABBCC\n")
#ggsave("article_profiling/fig_1b.pdf", fig_1b, height = 8, width = 15)

fig_1 <- wrap_plots(fig_1a, fig_1b, nrow = 2, heights = c(1, 0.85)) +
  plot_annotation(tag_levels = list(c("a", "b", "c", "d", "", "e", "f"))) &
  theme(plot.tag = element_text(face = "bold"))
ggsave("article_profiling/fig_1.pdf", fig_1, height = 16, width = 16)
#ggsave("article_profiling/fig_1.png", fig_1, height = 16, width = 16)
rm(fig_1a, fig_1b) #clean

#sup 1
fig_s1 <-  { plot_depth_top_phyla |
    (plot_depth_usgbs / plot_depth_spec + plot_layout(heights = c(0.85, 0.15))) } +
  plot_layout(ncol = 2, widths = c(0.25, 1)) +
  plot_annotation(tag_levels = list(c("a", "b", "c"))) &
  theme(plot.tag = element_text(size = 20, face = "bold"))
fig_s1[[2]][[1]] <- fig_s1[[2]][[1]] + theme(axis.title.y = element_text(vjust = 0))
fig_s1[[2]][[2]] <- fig_s1[[2]][[2]] + theme(axis.line.y.left = element_blank(),
                                             axis.ticks.y.left = element_blank())#,
                                             #axis.text.y = element_text(hjust = 2.2))
ggsave("article_profiling/fig_s1.pdf", fig_s1, height = 13, width = 22)
#ggsave("article_profiling/fig_s1.png", fig_s1, height = 13, width = 22)
```

@@

```{r processkos}
# Read output matrix from KOfamScan
kegg_annot <- read.csv("/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/analysis/kegg/merge_hits/kofamscan_hits.tsv", sep = "\t", check.names = FALSE)
colnames(kegg_annot)[2] <- "KO.definition"

@@

# Set definitions for pathways of interest
kegg_paths <- kegg_kos %>%
  #sulf: to keep sulfur, sulfate, sulfane, sulfide, sulfo_
  #soxr: SoxR 2Fe-2S cluster
  #nitrogen: a negative marker (ubiquous; ex: MerR family)
  dplyr::filter(grepl("iron|sulf|soxr|copper|nitrogen",
                      Definition, ignore.case = TRUE) == TRUE) %>%
  dplyr::select(KO, Definition)

>>>>

# Inspect and select sample S15
#filter(mags_qc, grepl("\\-S15\\.", bin)) %>% dplyr::select(2:3) %>% summary()
mag_samples <- dplyr::filter(gtdb_taxonomy, grepl("\\-S15\\.", genome))

@@

# Merge metadata and save to file
mags_metadata <- merge(mag_samples, smag_data, all = TRUE)
write.table(mags_metadata, "4-mags-smag/mags_metadata.tsv", sep = "\t", quote = FALSE, row.names = FALSE)

# Save parsed SMAG metadata
write.table(smag_data, "4-mags-smag/smag_data.tsv", sep = "\t", quote = FALSE, row.names = FALSE)

# Save metadata for sample S15
write.table(mag_samples, "4-mags-smag/mags_mining_S15.tsv", sep = "\t", quote = FALSE, row.names = FALSE)

@@

# Change ecosystem from "biomining" to "mining"
mag_samples <- mag_samples %>% dplyr::mutate(ecosystem = "Mining")
mags_metadata_bac <- mags_metadata_bac %>%
  dplyr::mutate(ecosystem = case_when(ecosystem == "Biomining" ~ "Mining", TRUE ~ ecosystem))

# Select top phyla and plot a histogram
mags_md_top_phy <- table(mags_metadata_bac$phylum) %>%
  sort(TRUE) %>%
  head(n = 40) %>% #top 40
  as.data.frame() %>%
  dplyr::rename(phy_top = Var1, freq = Freq)

plot_depth_top_phyla <- mags_md_top_phy %>%
  ggplot(aes(x = freq, y = reorder(phy_top, freq))) +
  geom_bar(stat = "identity", fill = "#459682") +
  geom_text(aes(label = freq), vjust = 0.5, hjust = -0.25, size = 3) +
  labs(x = "Number of MAGs", y = "Phylum") +
  scale_x_continuous(limits = c(0, 10000)) +
  theme_classic() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 17))
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_top_phyla.pdf", plot_depth_top_phyla, height = 8, width = 6)

# Filter metadata..
# Set colors for ecosystem
#mags_colors_eco <- colorRampPalette(brewer.pal(12, "Paired"))(length(unique(mags_metadata_bac$ecosystem)))
#mags_colors_eco <- sample(mags_colors_eco)
mags_colors_eco <- c("#1F78B4", "#FF7F00", "#B2DF8A", "#A6CEE3", "#6A3D9A",
                     "#33A02C", "#E31A1C", "#CAB2D6", "#FDBF6F", "#FB9A99")

# Which bacteria from mining sites are annotated as known?
mags_depth_spec_biom <- mags_metadata_bac %>%
  dplyr::select(mag_name, family, genus, species, ecosystem) %>%
  dplyr::filter(species != "" & ecosystem == "Mining")

@@

plot_depth_spec <- ggplot(mags_depth_spec) +
  geom_bar(aes(x = species, fill = phylum)) +
  labs(x = "", y = "Number of kSGBs from mining") +
  scale_fill_manual(values = "#78acc2") +
  coord_flip() +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = "top", #c(0.88, 0.12),
        strip.background = element_blank(),
        axis.text = element_text(size = 13),
        axis.title = element_text(size = 15),
        strip.text.x = element_text(size = 11)) +
  facet_grid(.~ecosystem)
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_biom_species.pdf", plot_depth_spec, height = 3, width = 9)

@@

#add column
mags_depth_spec_eco <- mags_depth_spec_eco %>%
  dplyr::mutate(mags = replace(mags, ecosystem == "Mining", 1)) %>%
  dplyr::mutate(ratio = count_eco / mags) #max 6.52

plot_depth_spec_eco <- mags_depth_spec_eco %>%
  ggplot(aes(x = reorder(ecosystem, ratio), y = ratio, fill = ecosystem)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(ratio, 2)), vjust = 0.5, hjust = -0.25, size = 3) +
  labs(x = "Ecosystem", y = "Number of kSGBs/samples") +
  scale_y_continuous(limits = c(0, 10)) +
  scale_fill_manual(values = mags_colors_eco) +
  coord_flip() +
  theme_classic() +
  theme(legend.position = "none",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_species_eco.pdf", plot_depth_spec_eco, height = 7, width = 6)

@@

plot_depth_sgbs <- mags_depth_sgbs[1:60, ] %>%
  ggplot(aes(x = reorder(phylum, tot), y = tot, fill = cond)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#dd7e8b", "#459682")) +
  scale_y_continuous(limits = c(0, 8000)) +
  labs(x = "Phylum", y = "Number of species") +
  coord_flip() +
  theme_classic() +
  theme(legend.position = c(0.85, 0.15),
        strip.placement = "outside",
        legend.text = element_text(size = 10),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))

@@

plot_depth_sgbs_p <- mags_depth_sgbs_p[1:60, ] %>%
  ggplot(aes(x = reorder(phylum, t_phy), y = tot_p, fill = cond_p)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#dd7e8b", "#459682")) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "", y = "% of species") +
  coord_flip() +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))
#merge
plot_depth_sgbs_both <- wrap_plots(plot_depth_sgbs, plot_depth_sgbs_p,
                                   nrow = 1, ncol = 2, widths = c(0.55, 0.5))
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_SGBs_phylum.pdf", plot_depth_sgbs_both, height = 7, width = 7)

# And unknown/known SGBs for mining?
mags_depth_sgbs_biom <- mags_metadata_bac %>%
  dplyr::filter(ecosystem == "Mining") %>%
  dplyr::group_by(phylum, species) %>%
  dplyr::summarise(n_spe = n()) %>%
  ungroup() %>%
  dplyr::group_by(phylum) %>%
  dplyr::mutate(t_phy = sum(n_spe)) %>%
  dplyr::filter(species == "") %>%
  dplyr::mutate(diff_k = t_phy - n_spe) %>%
  dplyr::mutate(perc_u = n_spe * 100 / t_phy) %>%
  dplyr::arrange(desc(t_phy)) %>%
  pivot_longer(cols = c(n_spe, diff_k), names_to = "cond", values_to = "tot")

plot_depth_sgbs_biom <- mags_depth_sgbs_biom %>%
  ggplot(aes(x = reorder(phylum, tot), y = tot, fill = cond)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#dd7e8b", "#459682")) +
  scale_y_continuous(limits = c(0, 25)) +
  labs(x = "Phylum", y = "Mining species") +
  coord_flip() +
  theme_classic() +
  theme(aspect.ratio = 0.875,
        legend.position = c(0.75, 0.3),
        strip.placement = "outside",
        legend.text = element_text(size = 10),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_SGBs_phylum_biom.pdf", plot_depth_sgbs_biom, height = 4, width = 6)

@@

plot_depth_usgbs <- mags_depth_usgbs_eco %>%
  ggplot(aes(x = num, y = reorder(phylum, num), fill = ecosystem)) +
  geom_col() +
  #geom_text(aes(label = num), vjust = 0.5, hjust = -0.5, size = 2.5) +
  labs(x = "Number of uSGBs", y = "Phylum") +
  scale_x_continuous(n.breaks = 3) +
  scale_fill_manual(values = mags_colors_eco) +
  theme_classic() +
  theme(legend.position = "none",
        strip.placement = "outside",
        strip.background = element_blank(),
        panel.border = element_blank(),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 18),
        strip.text.x = element_text(size = 12)) +
  facet_grid(.~ecosystem, scales = "free")
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_uSGBs_eco.pdf", plot_depth_usgbs, height = 11, width = 17)

# What about the annotated bins at genus level (kGGBs) from mining?
mags_depth_kggbs_biom <- mags_metadata_bac %>%
  dplyr::select(mag_name, mag_id, phylum, family, genus, ecosystem) %>%
  dplyr::filter(genus != "" & ecosystem == "Mining") #nrow=21

@@

plot_depth_kggbs_biom <- mags_depth_kggbs_biom %>%
  ggplot(aes(y = reorder(genus, norm), x = norm, fill = ecosystem)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_x_continuous(limits = c(0, 101), n.breaks = 5) +
  scale_fill_manual(values = mags_colors_eco) +
  labs(x = "% kGGBs", y = "kGGBs from mining") +
  theme_classic() +
  theme(legend.title = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_kGGBs_biom.pdf", plot_depth_kggbs_biom, height = 7, width = 6)

### --EXPLORE OTHER TAXA-- ###
##inspect unannotated family-level bins (uFGBs)? #1166
##what about novel bacteria at genus level?
#mags_depth_novel_gen <- mags_metadata_bac %>%
#  dplyr::filter(genus == "") %>%
#  dplyr::group_by(phylum, genus, ecosystem) %>%
#  dplyr::summarise(gen_unk = n()) %>%
#  #dplyr::arrange(desc(gen)) %>%
#  ungroup() %>%
#  dplyr::group_by(genus) %>%
#  dplyr::mutate(tot_gen = sum(gen_unk)) %>%
#  dplyr::mutate(perc = gen_unk * 100 / tot_gen)
#
##get unique classes of unknown genus by ecosystem
##Leptospirillia, Sulfobacillia
#mags_md_unk_gen2 <- mags_metadata_bac %>%
#  dplyr::filter(genus == "") %>%
#  dplyr::group_by(class, ecosystem) %>%
#  tally() %>%
#  #slice(n()) %>%
#  ungroup()
#unique(mags_md_unk_gen2[mags_md_unk_gen2[, "ecosystem"] == "Mining", "class"])
#
##unique genus from mining: Acidithrix, Ferrithrix, SCQM01
#unique(mag_samples[mag_samples[, "family"] != "", "genus"])
#mags_metadata_bac %>% filter(grepl("Acidithrix", genus, ignore.case = TRUE) == TRUE)
#
#plot_md_unk_cla <- ggplot(mags_md_unk_gen2[mags_md_unk_gen2$n > 10, ],
#                          aes(x = reorder(class, n), y = n, fill = ecosystem)) +
#  geom_bar(stat = "identity") +
#  labs(x = "Class", y = "Count") +
#  scale_y_continuous(limits = c(0, 250)) +
#  theme_minimal() +
#  coord_flip() +
#  facet_wrap(~ecosystem)
#ggsave("plot_md_unk_class_eco.pdf", plot_md_unk_cla, height = 7, width = 10)
#
##class of interest for acidic soils:
##Alphaproteobacteria, Gammaproteobacteria
##Acidobacteriae, Thermoleophilia, Sulfobacillia, Acidimicrobiia
#table(mags_metadata_bac[mags_metadata_bac[, "phylum"] == "Proteobacteria", "class"])
#table(mags_metadata_bac[mags_metadata_bac[, "class"] == "Acidimicrobiia", "order"])
#
##family of interest reported for Systemix [ref 2022]:
##Pseudomonadaceae (mayor), Flavobacteriaceae, Erwiniaceae
#mags_md_fams_acid <- mags_metadata_bac %>%
#  dplyr::filter(family == "Pseudomonadaceae") %>%
#  dplyr::group_by(ecosystem) %>%
#  dplyr::summarise(count = n())
```

@@

# Select sample S15 and export to file
mags_qc <- mags_qc %>% filter(grepl("Refined-S15", bin))
write.table(mags_qc, "article_profiling/2-statistics/mags_S15.txt", sep = "\t", quote = FALSE, row.names = FALSE)

@@

# Plot
plot_mags_qc <- mags_qcs %>%
  ggplot(aes(x = mags, y = fct_reorder(label, mags), fill = label)) +
  geom_col(position = "dodge2") +
  #geom_text(aes(label = mags), vjust = 0.5, hjust = -0.25, size = 3) +
  scale_fill_manual(values = c("#db7971", "#5ca650", "#859ac7")) +
  labs(x = "Number of MAGs", y = "MIMAG") +
  theme_classic() +
  theme(aspect.ratio = 0.875,
        legend.position = "none",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))
ggsave("article_profiling/2-statistics/plots/plot_biom_MAGs_QC.pdf", plot_mags_qc, height = 4, width = 6)
```

@@

##export to file parsed gtdb_taxonomy 1-taxonomy/gtdbtk_mags_S15.tsv
##checkm_parsed_bins.tsv 2-statistics/checkm_mags_S15.tsv

@@

## Figures 1-2 & supplementary 1

```{r fig1}
# Merge plots using patchwork
fig_1 <- wrap_plots(plot_mags_qc, plot_depth_sgbs_biom, plot_tree_s15_md,
                    design = "\nACCC\nBCCC", heights = c(0.85, 1)) +
  plot_annotation(tag_levels = list(c("a", "b", "c"))) &
  theme(plot.tag = element_text(size = 18, face = "bold"))
ggsave("article_profiling/fig_1.pdf", fig_1, height = 8, width = 15)
ggsave("article_profiling/fig_1.png", fig_1, height = 8, width = 15)

fig_2 <- wrap_plots(plot_depth_sgbs_both, plot_depth_spec_eco,
                    plot_depth_kggbs_biom, design = "\nAAABBCC\n") +
  plot_annotation(tag_levels = list(c("a", "", "b", "c"))) &
  theme(plot.tag = element_text(size = 18, face = "bold"))
ggsave("article_profiling/fig_2.pdf", fig_2, height = 8, width = 16)
ggsave("article_profiling/fig_2.png", fig_2, height = 8, width = 16)

#sup 1
fig_s1 <-  { plot_depth_top_phyla |
    (plot_depth_spec / plot_depth_usgbs + plot_layout(heights = c(0.15, 0.85))) } +
  plot_layout(ncol = 2, widths = c(0.25, 1)) +
  plot_annotation(tag_levels = list(c("a", "b", "c"))) &
  theme(plot.tag = element_text(size = 18, face = "bold"))
fig_s1[[2]][[1]] <- fig_s1[[2]][[1]] + theme(axis.line.y.left = element_blank(),
                                             axis.ticks.y.left = element_blank())#,
                                             #axis.text.y = element_text(hjust = 2.2))
fig_s1[[2]][[2]] <- fig_s1[[2]][[2]] + theme(axis.title.y = element_text(vjust = 0))
ggsave("article_profiling/fig_s1.pdf", fig_s1, height = 15, width = 24)
ggsave("article_profiling/fig_s1.png", fig_s1, height = 15, width = 24)
```

@@

```{r processkos}
# Read output matrix from KOfamScan
kegg_annot <- read.csv("article_profiling/4-mags-smag/kofamscan_hits.tsv", sep = "\t", check.names = FALSE)
colnames(kegg_annot)[2] <- "KO.definition"

@@

# Set definitions for pathways of interest
kegg_paths <- kegg_kos %>%
  #sulf: to keep sulfur, sulfate, sulfane, sulfide, sulfo_
  #soxr: SoxR 2Fe-2S cluster
  dplyr::filter(grepl("iron|sulf|soxr|copper", Definition, ignore.case = TRUE) == TRUE) %>%
  dplyr::select(KO, Definition)

## FUNCTIONS ##
# Function that returns the pathway based on definitions
fx_setKOpathways <- function(def) {
  if (grepl("copper", def, ignore.case = TRUE) == TRUE)
    return("Copper")
  if (grepl("iron", def, ignore.case = TRUE) == TRUE)
    return("Iron")
  if (grepl("sulf|sox", def, ignore.case = TRUE) == TRUE)
    return("Sulfur")
  else
    return("Other")
}
# Function to define types of metabolism
fx_setKOmetabtypes <- function(def) {
  #oxidation, reduction
  #fixation, hydrolysis, production, utilization, transport
  #sulfane involved in oxidative processes, as well as dehydrogenases
  if (grepl("oxida|oxidi|hydrog|hydrox|sulfane", def, ignore.case = TRUE) == TRUE)
    return("Oxidation")
  if (grepl("reduc", def, ignore.case = TRUE) == TRUE)
    return("Reduction")
  if (grepl("utili", def, ignore.case = TRUE) == TRUE)
    return("Utilization")
  if (grepl("hydrolase", def, ignore.case = TRUE) == TRUE)
    return("Hydrolysis")
  if (grepl("transferase", def, ignore.case = TRUE) == TRUE)
    return("Transferase")
  if (grepl("transport|binding", def, ignore.case = TRUE) == TRUE)
    return("Transport")
  if (grepl("resistance", def, ignore.case = TRUE) == TRUE)
    return("Resistance")
  if (grepl("carrier|chaperone", def, ignore.case = TRUE) == TRUE)
    return("Carrier")
  if (grepl("regulat", def, ignore.case = TRUE) == TRUE)
    return("Regulator")
  if (grepl("lyase", def, ignore.case = TRUE) == TRUE)
    return("Lyase")
  else
    return("Other")
}
# Function that returns the ecosystem given a MAG
fx_getMAGecosystem <- function(mag) {
  if (grepl("spades", mag, ignore.case = TRUE) == TRUE) {
    return("Mining")
  }
  else {
    mag_eco <- mags_metadata_bac %>%
      dplyr::mutate(bin = paste0(mag_name, ".", mag_id)) %>%
      dplyr::filter(bin == mag) %>%
      dplyr::select(ecosystem) %>%
      as.character()
    return(mag_eco)
  }
}
# Function that returns the genome size given a MAG
fx_getMAGlength <- function(mag) {
  mag_size <- mags_metadata_bac %>%
    dplyr::mutate(bin = paste0(mag_name, ".", mag_id)) %>%
    dplyr::filter(bin == mag) %>%
    dplyr::select(genome_size) %>%
    as.numeric()
  return(mag_size)
}
## END FUNCTIONS ##
